# Data transformation

In this data transformation part, we will separately introduce how we clean the three source of data.  

For the first data source [U.S. Diabetes Surveillance System](https://gis.cdc.gov/grasp/diabetes/DiabetesAtlas.html), we can see that for every CSV file we downloaded in this way, there are two useless rows on the top, one empty column at the end, and one row showing the website at the bottom. Initially, we use `skip = 2` to delete the top two useless rows. Then, we found in some files the column names starting with `X` in the front and containing symbols such as `...` inside. So, we change the column names to more concise and understandable labels. Finally, we can get the data in a nicer format. We use the same process to deal with all the data from [U.S. Diabetes Surveillance System](https://gis.cdc.gov/grasp/diabetes/DiabetesAtlas.html), such as Physical Inactivity, Obesity, and Diagnosed Diabetes files.  

The second data source is the [National Center for Health Statistics](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Questionnaire). The data downloaded from there are in `.XPT` form, so we use `read.xport` command under foreign library in R to read them.  

The third data source we used is [CDC WONDER](https://wonder.cdc.gov/ucd-icd10.html). And the data is in the form of `.txt` file, so we use the readr library and the `read_tsv` to read them into R. After importing the file into R, we notice there is an empty bote column on the left, and there is a huge amount of metadata at the bottom. We delect the empty note column and the rows contain metadata.

```{r}
# data from U.S. Diabetes Surveillance System
diabetes1 <- read.csv("DiabetesAtlasData-map.csv",skip = 2)
diabetes1 <- diabetes1[1:1026,]

diabetes_genders <- read.csv("DiabetesAtlasData_gender.csv",skip=2)
diabetes_age <- read.csv("DiabetesAtlasData_age.csv",skip=2)
diabetes_race <- read.csv("DiabetesAtlasData_race.csv",skip=2)
diabetes_education <- read.csv("DiabetesAtlasData_education.csv",skip=2)
diabetes_genders <- diabetes_genders[1:20,1:6]
diabetes_age <- diabetes_age[1:20,1:13]
diabetes_race <- diabetes_race[1:20,1:13]
diabetes_education <- diabetes_education[1:20,1:10]

library(magrittr)
age <- diabetes_age %>% 
  dplyr::select(Year,`X18.44...Percentage`,`X45.64...Percentage`,`X65.74...Percentage`,
       `X75....Percentage`)
colnames(age) <- c("Year","18-44","45-64","65-74","over74")

gender <- diabetes_genders %>% 
  dplyr::select(Year,`Male...Percentage`,`Female...Percentage`)
colnames(gender) <- c("Year","Male","Female")

race <- diabetes_race %>% 
  dplyr::select(Year,`Non.Hispanic.White...Percentage`,`Non.Hispanic.Black...Percentage`,
                `Non.Hispanic.Asian...Percentage`,`Hispanic...Percentage`)
colnames(race) <- c("Year","White","Black","Asian","Hispanic")

education <- diabetes_education %>% 
  dplyr::select(Year,`X..High.School...Percentage`,`High.School...Percentage`,
                `X..High.School...Percentage.1`)
colnames(education) <- c("Year","< High School","High School",
                      "> High School")

DiagnosedDiabetes_new <- read.csv("DiabetesAtlasData_Diagnosed Diabetes.csv",skip = 2)
Obesity_new <- read.csv("DiabetesAtlasData _Obesity.csv",skip = 2)
PhysicalInactivity_new <- read.csv("DiabetesAtlasData_Physical_Inactivity.csv",skip = 2)
HypertensionAwareness_new <- read.csv("DiabetesAtlasData_Hypertension_Awareness.csv",skip=2)
HighCholesterolAwareness_new <- read.csv("DiabetesAtlasData_High_Cholesterol_Awareness.csv",skip=2)

DiagnosedDiabetes<- read.csv("DiagnosedDiabetes.csv",skip = 2)
Obesity <- read.csv("Obesity.csv",skip = 2)
PhysicalActivity<- read.csv("PhysicalActivity.csv",skip = 2)
```


```{r}
# questionnaire result from National Center for Health Statistics
library(foreign)
questionnaire18 <- read.xport("17_18.XPT")
```

```{r}
# data from CDC WONDER
library(readr)
death <- read_tsv("Underlying Cause of Death, 1999-2020.txt")
death <- death[1:10383,2:14]
death_year <- read_tsv("Underlying Cause of Death, 1999-2020_year.txt")
death_year <- death_year[1:1122,2:8]
death_month <- read_tsv("Underlying Cause of Death, 1999-2020_month.txt")
death_month <- death_month[1:13067,2:8]
```



